Competing Bandits: \\ The Perils of Exploration under Competition

### 2018-11-04 paper/content/introduction.tex
intro

Even though, under our model, consumers are best off when there is a ``temporary" monopoly, a recent subject of debate in policy circles is whether the data advantage alone poses a sufficiently powerful barrier to entry to warrant anti-trust interest. While \cite{varian2018artificial} and \cite{bajari2018impact} argue that, from a statistical perspective, data has decreasing returns to scale and therefore does not serve as a significant competitive advantage, they focus on the problem in isolation. We argue that, to reason about this question, one needs to consider what the impact of better data is when data collection has competitive consequences, especially when firms need to learn while competing. Therefore, we run additional experiments that attempt to isolate the impact of the data advantage and reputation advantage. We find that the data advantage alone still leads to a significant incumbency advantage. Additionally, the data advantage tends to be stronger when the incumbent commits to a more advanced bandit algorithm. This highlights that, when thinking about data as a barrier to entry, data quality and not just quantity matters. In particular, the data advantage may be larger and more relevant when the incumbent possesses data that it gets via exploration of potentially sub-optimal actions that would be costly for the entrant to gather upon entry.}

>>>>

Even though, under our model, consumers are best off when there is a ``temporary" monopoly, a recent subject of debate in policy circles is whether the data advantage alone poses a sufficiently powerful barrier to entry to warrant anti-trust interest. While \cite{varian2018artificial} and \cite{bajari2018impact} argue that, from a statistical perspective, data has decreasing returns to scale and therefore does not serve as a significant competitive advantage, they focus on the problem in isolation. We argue that, to reason about this question, one needs to consider what the impact of better data is when data collection has competitive consequences, especially when firms need to learn while competing. Therefore, we run additional experiments that attempt to isolate the impact of the data advantage and reputation advantage. We find that the data advantage alone still leads to a significant incumbency advantage. Additionally, the data advantage tends to be stronger when the incumbent commits to a more advanced bandit algorithm. This highlights that, when thinking about data as a barrier to entry, data quality and not just quantity matters. In particular, the data advantage may be larger and more relevant when the incumbent possesses data that it gets via exploration of potentially sub-optimal actions that would be costly for the entrant to gather upon entry.}\swcomment{this is quite confusing. let's talk}
